Simulating Generalized Fading Channels in Python

### 2018-09-13 paper/manuscript.tex
minor corrections

\begin{abstract}
    We present a novel and well tested Python-based library for simulating and computing
    generalized fading channels, considering $\alpha\text{-}\mu$, $\kappa\text{-}\mu$, and
$\eta\text{-}\mu$ distributions. We describe the
    applicability of this library using a few communications channels scenarios
    impaired by generalized fading, namely, spectrum sensing and bit error
    rate computation. The development of the library is open source and its code along
    with examples are available at
    \texttt{http://github.com/mirca/maoud}.
\end{abstract}

@@

The study of modern wireless communications systems heavily relies on fading channel simulation.
By fading channel simulation, we refer to the generation of samples from a probability distribution
that resembles the effects and impairments caused by real communications fading channels on the
transmitted signal.

@@

In this paper, we present an open source Python package
for generation of samples following the $\alpha\text{-}\mu$, $\kappa\text{-}\mu$, and
$\eta\text{-}\mu$ distributions. The usefullness of ths package is illustrated
through well known examples involving spectrum sensing and bit error rate (BER) computation.

Section~\ref{rejec} presents the well known rejection method which is used as
the basis for the proposed implementation. Section~\ref{sensing} and~\ref{ber}
demonstrates the features of the proposed package in spectrum sensing over complex
generalized fading channels and bit error rate computation on $\alpha$-$\mu$ fading.
Section~\ref{conclusions} states the final remarks.

@@

Basically, the steps of the rejection algorithm are\cite{ross}:
\begin{enumerate}
	\item generate a sample $y$ from the probability density $g$;
    \item generate a sample $u$ from a uniform density with support $(0, 1)$;
	\item check $u \leq \frac{f(y)}{cg(y)}$; if true then accept $y$; otherwise, return to step 1.
\end{enumerate}

@@

As a result, the probabilities of false alarm and miss detection can
be written as
\begin{align}
    \mathbb{P}\left(d_\delta(\tilde{y}) = 1 | H_0\right) &=
        1 -  P(\delta | H_0) = 1 - \gamma\left(n, \frac{\delta}{\sigma^2}\right),\label{eq:pf} \\
    \mathbb{P}\left(d_\delta(\tilde{y}) = 0 | H_1\right) &= \mathbb{E}_{h}\left(P(\delta| h, H_1)\right)\nonumber\\
    &= \int_{-\infty}^{+\infty} P(\delta | h, H_1)p(h)\;\mathrm{d}h,\label{eq:pd}
\end{align}
in which $\gamma$ is the regularized lower incomplete Gamma function, $p(h)$ is the pdf of the fading,
and $\delta$ is the energy detection threshold.

@@

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/spectrum_sensing}
    \caption{Receiver Operating Characteristic for the energy detector in $\alpha$-$\mu$
             fading channel with $\alpha=2$ and $\mu=1$.
             }
    \label{fig:spectrum-sensing}
\end{figure}

@@

Consider the system
\begin{align}
    y = hs + w
\end{align}
in which $s \in \{0, a\}$, $a \in \mathbb{R}_{+}$, is a transmitted signal,
$h$ is an $\alpha$-$\mu$ random variable and $\bm{w}$ is a Gaussian random variable
with zero mean vector and variance $\sigma^2$, and $y$ is the received signal.

@@

Further, assume that the decoded bit $\hat{y}$ is estimated using the minimum distance decoding
rule, i.e.,
\begin{align}
    \hat{y} = \mathbb{I}(|y - a|^2 < |y|^2)
\end{align}
therefore
\begin{align}
    \mathbb{P}\left(\hat{y} = 1 | s = 0\right) & = \mathbb{P}\left(|w - a|^2 - |w|^2 < 0\right)\nonumber\\
    & = \mathbb{P}\left(w > \frac{a}{2}\right) = 1 - \Phi\left(\frac{a}{2\sigma}\right)
\end{align}
and likewise
\begin{align}
    \mathbb{P}\left(\hat{y} = 0 | s = a\right) & = \mathbb{P}\left(ha + w < \frac{a}{2}\right) \nonumber\\
    & = \mathbb{E}_h\left(\Phi\left(\frac{a(1 - 2h)}{2\sigma}\right)\right).
\end{align}

Fig.~\ref{fig:ber} shows the expression of the BER as a function of the signal-to-noise ratio.
As we can see, the simulation presents an excellent agreement against theorectical results.

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/ber}
    \caption{BER \textit{versus} SNR.}
    \label{fig:ber}
\end{figure}

@@

>>>>

\begin{abstract}
    We present a novel and well tested Python-based library for simulating and
    computing generalized fading channels considering $\alpha\text{-}\mu$,
    $\kappa\text{-}\mu$, and $\eta\text{-}\mu$ distributions. We describe the
    applicability of this library through common communications engineering
    use cases such as spectrum sensing and bit error rate computation.
    The development of the library is open source and its code, along with
    examples, are available at
    \texttt{http://github.com/mirca/maoud}.
\end{abstract}

@@

The study of modern wireless communications systems heavily relies on fading channel simulation.
By fading channel simulation, we refer to the generation of samples from a probability distribution
that resembles the effects and impairments caused by real communications channels on the
transmitted signal.

@@

In this paper, we present an open source Python package
for generation of samples following the $\alpha\text{-}\mu$, $\kappa\text{-}\mu$, and
$\eta\text{-}\mu$ distributions. The usefullness of this package is illustrated
through well known examples involving spectrum sensing and bit error rate
(BER) computation.

@@

Basically, the steps of the rejection algorithm are~\cite{ross}:
\begin{enumerate}
	\item generate a sample $y$ from the probability density $g$;
    \item generate a sample $u$ from a uniform density with support $(0, 1)$;
	\item check $u \leq \frac{f(y)}{cg(y)}$; if true then accept $y$; otherwise, return to step 1.
\end{enumerate}

@@

As a result, the probabilities of false alarm and miss detection can
be written as
\begin{align}
    \mathbb{P}\left(d_\delta(\tilde{y}) = 1 | H_0\right) &=
        1 -  P(\delta | H_0) = 1 - \gamma\left(n, \frac{\delta}{\sigma^2}\right),\label{eq:pf} \\
    \mathbb{P}\left(d_\delta(\tilde{y}) = 0 | H_1\right) & = \int_{-\infty}^{+\infty} P(\delta | h, H_1)p(h)\;\mathrm{d}h,\label{eq:pd}
\end{align}
in which $\gamma$ is the regularized lower incomplete Gamma function, $p(h)$ is the pdf of the fading,
and $\delta$ is the energy detection threshold.

@@

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/spectrum_sensing}
    \caption{Receiver Operating Characteristic for the energy detector in $\alpha$-$\mu$
             channel with $\alpha=2$ and $\mu=1$.
             }
    \label{fig:spectrum-sensing}
\end{figure}

@@

Consider the system
\begin{align}
    y = hs + w
\end{align}
in which $s \in \{0, a\}$, $a \in \mathbb{R}_{+}$, is a transmitted signal,
$h$ is an $\alpha$-$\mu$ random variable and $w$ is a Gaussian random variable
with zero mean vector and variance $\sigma^2$, and $y$ is the received signal.

@@

Further, assume that the decoded bit $\hat{y}$ is estimated using the minimum distance decoding
rule, i.e.,
\begin{align}
    \hat{y} = \mathbb{I}(|y - a|^2 < |y|^2)
\end{align}
therefore
\begin{align}
    \mathbb{P}\left(\hat{y} = 1 | s = 0\right) & = \mathbb{P}\left(|w - a|^2 - |w|^2 < 0\right) = 1 - \Phi\left(\frac{a}{2\sigma}\right)
\end{align}
and likewise
\begin{align}
    \mathbb{P}\left(\hat{y} = 0 | s = a\right) & = \mathbb{P}\left(ha + w < \frac{a}{2}\right) = \mathbb{E}_h\left(\Phi\left(\frac{a(1 - 2h)}{2\sigma}\right)\right).
\end{align}

In Fig.~\ref{fig:ber}, we show the expression of the BER as a function of the signal-to-noise ratio,
for the system under consideration with $\alpha=2$ and $\mu = 5$.
As we can see, the simulation presents an excellent agreement against theorectical results.
\begin{figure}[!htb]
    \centering
    \includegraphics{figures/ber}
    \caption{BER as a function of the SNR for $\alpha=2$ and $\mu = 5$}
    \label{fig:ber}
\end{figure}

@@

Finally, all the simulations were done in a personal laptop,
and the generation of $10^6$ samples from any of the generalized fading
distributions took on average 10 seconds.
